Applications of Technology and Large Data in Statistics Education and Statistical Graphics

### 2015-05-27 DefenseSlides/DefenseSlidesCommitteeHour.Rnw
Subscripts and citation start
Reference file not working, directly code authors for slides
Edits to slides 1-13 and added motivation slide

\documentclass{beamer} 
% \usepackage{graphicx}
\usepackage{graphics}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{bbm}

\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx,float,wrapfig,subfig,tabularx,ulem}
\usepackage{color}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{mathrsfs}

@@

@@

Change Point Rendering
  \begin{itemize}
    \item Open Points
    \item Alpha-Blending
    \item Generalized Scatterplots
  \end{itemize} \pause

Plotting Binned Aggregations
  \begin{itemize}
    \item Sunflower Plots
    \item Bubble Plots
    \item Binned Scatterplots
  \end{itemize}

@@

\begin{frame}
    \frametitle{Binning Functions}

Goal for Binned Scatterplot:
  \begin{itemize}
    \item Reduced Binned Data Triples $(x^\ast, y^\ast, c)$
    \item $(x^\ast,y^\ast)$ = bin center
    \item $c$ = bin frequency count
  \end{itemize}

Need binning function to uniquely map points to bins
  \begin{itemize}
    \item $b(.) : (x,y) \rightarrow (x^\ast,y^\ast)$
  \end{itemize}

@@

Rectangular binning bins each dimension univariately
  \begin{itemize}
    \item $b(x_i,y_i) = (b_X(x_i),b_Y(y_i))$
    \item Forms rectangular binned grid
  \end{itemize}
\pause

Generally $b_X(.) : x_i \rightarrow x^\ast_j$ is defined as 
%
\begin{eqnarray}\label{rectbin}
b_X(x_i) = \left\{\begin{array}{ll} 
  x^\ast_{1} &\text{ for all } x_i = \beta_{0} \\
  x^\ast_j & \text{ for all } x_i \in (\beta_{j-1} , \beta_j] 
  \end{array}\right\}
\end{eqnarray}

where $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$ are sequence of 
$J$ adjacent intervals spanning range of X data.

@@

How do we discretize the frequency shading?
  \begin{itemize}
    \item Add second stage of binning
    \item Use one dimensional binning algorithms on bin counts, $c_\ell$
    \item Shade ordinally based on binned frequency groups
    \item May transform counts before frequency binning \\ (i.e. log counts)
  \end{itemize}

@@

Log Frequency Loss
  \begin{itemize}
    \item If bin frequencies, $c_\ell$, are log transformed 
    \item binned log frequencies, $b_C(\log(c_\ell))$
  \end{itemize} 
  %
\begin{eqnarray}\label{logfrequency}
L^{\log F} = \sum_{\ell \in \mathscr{L}^\ast} (\log(c_\ell) - b_c(\log(c_\ell)))^2
\end{eqnarray}

@@

>>>>

\documentclass{beamer} 
% \usepackage{graphicx}
\usepackage{graphics}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{bbm}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx,float,wrapfig,subfig,tabularx,ulem}
\usepackage{color}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{mathrsfs}

\bibliographystyle{asa}

@@

%------------------------------------------------
\begin{frame}
    \frametitle{Motivation}

Why use binned scatterplots?
  \begin{itemize}
    \item Cannot effectively display large data with the traditional scatterplot
    \item Make information 'lost in translation' from visually rendering and reading the bivariate relationship explicit
  \end{itemize}

Primary goals:
  \begin{itemize}
    \item Introduce spatial and frequency loss
    \item Investigate binning properties
  \end{itemize}

@@

Change Point Rendering
  \begin{itemize}
    \item Open Points (Tukey, 1977)
    \item Alpha-Blending (Unwin, Theus \& Hofmann, 2006)
    \item Generalized Scatterplots (Keim, Hao, Dayal, Janetzko \& Bak, 2010)
  \end{itemize} \pause

Plotting Binned Aggregations
  \begin{itemize}
    \item Sunflower Plots (Cleveland and McGill, 1984)
    \item Bubble Plots (Playfair, 1786)
    \item Binned Scatterplots (Unwin, Theus \& Hofmann, 2006)
  \end{itemize}

@@

\begin{frame}
    \frametitle{Constructing Binned Scatterplots}

Reduce binned data triples:
  \begin{itemize}
    \item $(x^\ast_\ell, y^\ast_\ell, c_\ell)$ for all bins $\ell \in \{1, \dots, \mathscr{L} \}$
    \item $(x^\ast_\ell,y^\ast_\ell)$ = center of bin $\ell$
    \item $c_\ell$ = frequency count for bin $\ell$
    \item Triples are \textit{sufficient} for creating binned scatterplot
  \end{itemize}

\vspace{.1in}

Binning functions uniquely map data pairs to bin centers
  \begin{itemize}
    \item $b(.) : (x_i,y_i) \rightarrow (x^\ast_\ell,y^\ast_\ell)$ 
    \item Where $(x_i,y_i)$ are data pairs for observations $i \in \{1,\dots,n\}$
  \end{itemize}

@@

Rectangular binning functions bin each dimension univariately
  \begin{itemize}
    \item $b(x_i,y_i) = (b_X(x_i),b_Y(y_i))$
    \item Forms rectangular binned grid
    \item Computationally efficient (Wickham, 2013)
    \item Can generalize to higher dimension easily
  \end{itemize}

Generally, $b_X(.) : x_i \rightarrow x^\ast_j$ is defined as 
%
\begin{eqnarray}\label{rectbin}
b_X(x_i) = \left\{\begin{array}{ll} 
  x^\ast_{1} &\text{ for all } x_i = \beta_{0} \\
  x^\ast_j & \text{ for all } x_i \in (\beta_{j-1} , \beta_j] 
  \end{array}\right\}
\end{eqnarray}

where $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$ are sequence of 
$J$ adjacent intervals coverting range of the data. (Scott, 1979)

\vspace{.1in}

Parameters must be specified such that: $\beta_0 \le x_{(0)} \le x_{(n)} \le \beta_J$
where $x_{(1)}$ and $x_{(n)}$ are the minimum and maximum data values

\vspace{.1in}

@@

How do we discretize the frequency shading?
  \begin{itemize}
    \item Add second stage of binning
    \item Use one dimensional binning algorithms on bin counts
    \item Shade ordinally based on binned frequency groups
    \item May transform counts before frequency binning \\ (i.e. log counts)
  \end{itemize}

@@

Log Frequency Loss
  \begin{itemize}
    \item If bin frequencies, $c_\ell$, are log transformed 
    \item binned log frequencies, $b_C(\log(c_\ell))$
  \end{itemize} 
  %
\begin{eqnarray}\label{logfrequency}
L^{\log F} = \sum_{\ell=1}^{\mathscr{L}} (\log(c_\ell) - b_c(\log(c_\ell)))^2
\end{eqnarray}

@@

% \bibliography{references}
